---
title: "research part 2"
author: "YZ"
date: "2023-11-04"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r data load}
library(MASS)
library(ggplot2)
library(readxl)
library(rcarbon)
library(gridExtra)
library(dplyr)
library(cowplot)
df_path = "hunter data.xlsx"
df <- read_excel(df_path, sheet = "orginial data")
df$sd_c14 = (df$max - df$min) / sqrt(12)
# calculate r和p
df <- df %>%
  mutate(r = (`Nb years`^2) / (`std err years`^2 - `Nb years`),
         p = r / (r + `Nb years`)) %>%
  filter(!(r < 0 | p < 0 | is.nan(r) | is.nan(p)))

x_max_values <- numeric(nrow(df))

# Loop to calculate PMF, CDF and peak value for each position
for (i in 1:nrow(df)) {
  # get mean_val和sd_val
  r <- df$`r`[i]
  p <- df$`p`[i]
  
  x_values <- 0:3000

  # calculate CDF
  cdf_values <- pnbinom(x_values, size = r, prob = p)
  
  # calculate PMF
  pmf_values <- dnbinom(x_values, size = r, prob = p)
  
  # find min for CDF over 0.99
  x_max <- min(x_values[cdf_values > 0.99])
  x_max_values[i] <- x_max
  
  # find PMF peak
  peak_x <- x_values[which.max(pmf_values)]
  peak_y <- max(pmf_values)
  
  # Create PMF diagram
  plot_pmf <- ggplot(data.frame(x_values, pmf_values), aes(x = x_values, y = pmf_values)) +
    geom_line(color = 'blue') +
    geom_point(aes(x = peak_x, y = peak_y), color = 'red', size = 3) +
    labs(title = df$group[i], x = "Generation Number", y = "Probability Mass Function Value")
  print(plot_pmf)
  # Create CDF diagram
  plot_cdf = ggplot(data.frame(x_values, cdf_values), aes(x = x_values, y = cdf_values)) +
    geom_line(color = 'red') +
    geom_vline(xintercept = x_max, linetype="dashed") +
    labs(title = df$group[i], x = 'Generation Number', y = 'Cumulative Distribution Function Value')
  print(plot_cdf)
}

df$x_max <- x_max_values
```

```{r}
library(openxlsx)
library(MASS)
library(ggplot2)
library(readxl)
library(rcarbon)
library(gridExtra)
library(dplyr)
library(cowplot)

for (i in 1:nrow(df)) {
    # function for C in rcarbon
  get_probability_c <- function(value) {
    index <- which(data_points == value)
    if (length(index) > 0) {
      return(probabilities[index])
    } else {
      return(0)  #Return 0 if the value isn't in data_points
    }
  }
  
  # Function to get the probability from the negative binomial distribution
  get_probability_from_negbin <- function(value) {
    return(dnbinom(x = value, size = r, prob = 1-p))
  }
  
  r = df$r[i]
  p = df$p[i]
  # 计算每个位置的mean_val和sd_val
  mean_val <- (df$max[i] + df$min[i]) / 2
  sd_val <- df$sd_c14[i]
  x_max = df$x_max[i]
  # 使用rcarbon包的calibrate函数进行校准
  x <- calibrate(x=mean_val, errors=sd_val, calCurves='intcal20')
  
  # 创建绘图
  plot(x, HPD=TRUE, credMass=0.95)
  
  # 获取数据点和概率值
  data_points <- rev(c(x$grids$'1'$calBP))
  probabilities <- rev(c(x$grids$'1'$PrDens))
    # Store the inside_p and Sum_p values for each Time
  Time_values <- c()
  inside_p_values <- c()
  
  generation_max= x_max+max(data_points)

    
  for (generation in min(data_points):generation_max) {
    inside_p = 0
    
    for (data_point in data_points) {
      generation_c = generation-data_point
      if (generation_c > 0) {
        negbin = round(generation_c)
        p_c = get_probability_from_negbin(negbin) # Make sure r and p are properly defined
        p_data_point = get_probability_c(data_point)
        inside_p = inside_p + p_data_point * p_c
      }
    }
    new_time_value = generation
    if(new_time_value %in% Time_values) {
      index <- which(Time_values == new_time_value)
      # Accumulate the new value at the corresponding position in inside_p_values
      inside_p_values[index] <- inside_p_values[index] + inside_p
    } else {
      # Add the new elements (probability)
      Time_values <- c(Time_values, new_time_value)
      inside_p_values <- c(inside_p_values, inside_p)
      }
  }
    
  # Sort Time_values and inside_p_values
  sorted_indexes <- order(Time_values)
  Time_values_sorted <- Time_values[sorted_indexes]
  inside_p_values_sorted <- inside_p_values[sorted_indexes]
  excel_name = df$group[i]
    # Define the file name for the Excel file
  excel_file_name <- paste0("G:\\Aoife\\research p2\\", excel_name, ".xlsx")
  print(excel_name)
  # Create a new workbook
  wb <- createWorkbook()

  # Add a new worksheet
  addWorksheet(wb, "Data")

  # Write Time_values_sorted and inside_p_values_sorted to the worksheet
  writeData(wb, "Data", Time_values_sorted, startCol = 1, startRow = 1, colNames = FALSE)
  writeData(wb, "Data", inside_p_values_sorted, startCol = 2, startRow = 1, colNames = FALSE)

  # Save the workbook
  saveWorkbook(wb, excel_file_name, overwrite = TRUE)
  
  # Define the file name based on the group and the location to save it
  file_name <- paste0("G:\\Aoife\\research p2\\", df$group[i], ".png")
  
  # Open a PNG device
  png(filename=file_name, width=1200, height=1000)
  # Adjust label and axis font sizes
  cex_lab <- 1.7  
  cex_axis <- 1.7 
  cex_main <- 2.0 
  par(mar=c(5.1, 5.1, 4.1, 2.1))
  # Plot the probability values for Time with adjusted font sizes
  plot(Time_values_sorted, inside_p_values_sorted, type='l', 
       col='blue', xlab='Time (BC))', ylab='Probability', 
       main=paste('Probability values for C at ', df$group[i], ' in Time frame'), 
       xlim=rev(range(Time_values)), 
       ylim=c(min(inside_p_values_sorted), max(inside_p_values_sorted) * 1.1), 
       cex.lab=cex_lab, cex.axis=cex_axis, cex.main=cex_main)

  
  # Identify the peaks in the inside_p_values_sorted
  peak_indices <- which(diff(sign(diff(inside_p_values_sorted))) == -2)
  
  # Ensure we only have the top 3 peaks
  if (length(peak_indices) > 3) {
    peak_values <- inside_p_values_sorted[peak_indices]
    peak_indices <- peak_indices[order(peak_values, decreasing = TRUE)[1:3]]
  }
  
  # Plot and label the peaks with reduced size and adjusted position


